Minimum Regret Search for Single- and Multi-Task Optimization

### 2016-06-07 2016_ICML/presentation/main.tex
Working on presentation

\newcommand{\newblock}{}

% Author, Title, etc.
\title[EWRL 2013]{Learning Skill Templates for Parameterized Tasks}
\author[Metzen \& Fabisch]{Jan Hendrik Metzen and Alexander Fabisch}
\institute[University Bremen]{AG Robotik, University Bremen\\ Project BesMan \url{http://robotik.dfki-bremen.de/de/forschung/projekte/besman.html}}
\date[05-09/08/13]{EWRL, August 05-09, 2013}

@@

\frame{\titlepage
       \bibliographystyle{abbrv}
       \nobibliography{../references}}

\begin{frame}{Motivation: Learn Skills for Robotic Mobile Manipulation}
   \begin{center}
   \movie[width=8cm,height=6cm,externalviewer]
             {\includegraphics[width=8cm,height=6cm]{PA10_Throwing_Ball}}{targets_short.avi} \\
   Example: learn to throw a ball at different targets
   \end{center}
\end{frame}

\begin{frame}{Multi-Task Reinforcement Learning}
  Given:
  \begin{itemize}
    \item Class of RL problems parameterized by $\taskparam \in T \subseteq \mathbb{R}^n$
    \item Task distribution $P(\taskparam)$ {\tiny [in this talk uniform over $T$]}
    \item Skill represented by a parameterized policy with parameter vector $\skillparam \in
\mathbb{R}^m$
    \item Means for determining return $J(\skillparam, \taskparam)$ of the skill parametrized by $\skillparam$ in task $\taskparam$
  \end{itemize}
  \pause
  Parameterized skill \footcite{da_silva_learning_2012}:
  \begin{itemize}
    \item Learn a function $\hat{\Theta}: \taskparam \mapsto \skillparam_\taskparam$ from task vector $\taskparam$ to a skill vector $\skillparam_\taskparam$
    \item Objective: $\hat{\Theta}$ similar to $\Theta^* = \arg\max_\Theta \int P(\taskparam)J(\Theta(\taskparam),\taskparam)d\taskparam$
    \item Typically: $\hat{\Theta}(\taskparam) \neq \arg\max_\skillparam J(\skillparam,\taskparam)$
  \end{itemize}
  \pause
  Our approach: give more than point estimate $\hat{\Theta}(\taskparam)$
\end{frame}

\begin{frame}{Learning Skill Templates}
  Given:
  \begin{itemize}
    \item Training set from $K$ source tasks $E = \left\{(\taskparam_i, \skillparam_{\taskparam_i}) \vert i=1,\dots,K\right\}$ with $J(\skillparam_{\taskparam_i},\taskparam_i) \approx J(\skillparam_{\taskparam_i}^*,\taskparam_i)$
    \item $E$ can be generated using RL (e.g., direct policy search) in each of the tasks
  \end{itemize}
 \pause
 Skill Templates
 \begin{itemize}
    \item \emph{Goal}: Transfer procedural knowledge contained in $E$ to novel, $(K+1)$st tasks $\tau$
    \item \emph{Idea}: learn to initialize policy search for new task such that close-to-optimal policies can be learned more sample-efficiently
    \item Skill template: $\psi: \taskparam \mapsto \mathcal{N}(\Theta(\taskparam), \Omega(\taskparam))$
    \item Normal distribution over policy parameter space, where
    \begin{itemize}
      \item $\Theta(\taskparam) \in \mathbb{R}^m$ determines starting point for policy search
      \item $\Omega(\taskparam) \in \mathbb{R}^{m\times m}$ controls initial exploration
    \end{itemize}
    \item Skill template learned from $E$
  \end{itemize}

@@

\begin{frame}{Approach}
  \begin{itemize}
    \item Represent skills using dynamical systems movement primitives (DMPs)\footcite{Ijspeert2013}\footcite{mulling_learning_2012}
    \item Learn source task policies using policy search based on CMA-ES \footcite{hansen_completely_2001}
    \item Learn $\Theta: \taskparam \mapsto \skillparam_\taskparam$ from E using Gaussian Process Regression (GPR) \footcite{rasmussen_gaussian_2006}
    \item Restrict $\Omega$ to the diagonal case {\tiny [for now]}:
    \begin{enumerate}
      \item Weight-proportional exploration: $\Omega(\taskparam)_{ii}$ proportional to $\vert\Theta(\taskparam)_i\vert$
      \item Uncertainty exploration: $\Omega(\taskparam)_{ii}$ proportional to uncertainty in GPR's prediction of $\Theta_i$
    \end{enumerate}
    \item Normalize mean eigenvalue of $\Omega(\taskparam)$
  \end{itemize}
\end{frame}

\begin{frame}{Test domain}
  \begin{itemize}
    \item Learn to throw a ball with a simulated PA-$10$
    \item $\taskparam \in \mathbb{R}^2$: target position on floor
    \item $\skillparam \in \mathbb{R}^{96}$: DMP parameters controlling robot in joint space ($8$ joints):
    \begin{itemize}
      \item Joint's trajectory ($10$ parameters per joint)
      \item Joint's final position and velocity ($2$ parameters per joint)
    \end{itemize}
    \item $J(\skillparam,\taskparam)$: Negative of squared distance of thrown ball from target
  \end{itemize}
  \begin{center}
    \includegraphics[width=.6\linewidth]{PA10_Throwing_Ball}
  \end{center}
\end{frame}

\begin{frame}{Illustration: Ball Throwing (1/2)}
   \begin{center}
   Error of $\Theta(\taskparam)$ vs.\,Number of Source Tasks
   \end{center}
   \begin{center}
         \only<1>{\includegraphics[width=\linewidth]
                   {plots/prediction_error_00}}
         \only<2>{\includegraphics[width=\linewidth]
                   {plots/prediction_error_01}}
         \only<3>{\includegraphics[width=\linewidth]
                   {plots/prediction_error_03}}
         \only<4>{\includegraphics[width=\linewidth]
                   {plots/prediction_error_05}}
         \only<5>{\includegraphics[width=\linewidth]
                   {plots/prediction_error_09}}
   \end{center}
\end{frame}

\begin{frame}{Illustration: Ball Throwing (2/2)}
   \begin{center}
   Error during Learning (Tabula Rasa vs.\,Skill Template (Uncertainty))
   \end{center}
   \begin{center}
         \only<1>{\includegraphics[width=\linewidth]
                   {plots/error_during_learning_00}}
         \only<2>{\includegraphics[width=\linewidth]
                   {plots/error_during_learning_01}}
         \only<3>{\includegraphics[width=\linewidth]
                   {plots/error_during_learning_02}}
         \only<4>{\includegraphics[width=\linewidth]
                   {plots/error_during_learning_03}}
         \only<5>{\includegraphics[width=\linewidth]
                   {plots/error_during_learning_04}}
         \only<6>{\includegraphics[width=\linewidth]
                   {plots/error_during_learning_05}}
   \end{center}
\end{frame}

\begin{frame}{Results: Ball Throwing}
   \begin{center}
   Learning Curve (Tabula Rasa vs.\,ST with different $\Omega$)
   \end{center}
   \begin{center}
         \includegraphics[width=.8\linewidth]{plots/learning_curve}\\
         Initial exploration covariance matrix matters (even with CMA-ES)
   \end{center}
\end{frame}

\begin{frame}{Outlook and Discussion}
  Open questions and future work:
  \begin{itemize}
    \item Correlated skill parameter dimensions: how to learn non-diagonal $\Omega \in \mathbb{R}^{m \times m}$ from small set of source tasks?
    \item Off-task learning: how to learn from $\skillparam_{\taskparam}$ with $J(\skillparam_{\taskparam},\taskparam) \ll J(\skillparam_{\taskparam}^*,\taskparam)$?
    \item Best way to handle discontinuities in mapping $\Theta$ (like fore- vs.\,backhand)?
    \item Learn $\Theta$ using (non-parametric) regression or (parametric) policy search?
    \pause
    \item Evaluation on humanoid robot AILA (simulated and real)
  \end{itemize}

  \begin{center}
    \includegraphics[width=.39\linewidth]{AILA}
  \end{center}
  \pause
  \begin{center}
   Thank you for your attention!
   Do you have questions, comments, or ideas?
   \end{center}

>>>>

%\newcommand{\newblock}{}

% Author, Title, etc.
\title[ICML 2016]{Minimum Regret Search for Single- and Multi-Task Optimization}
\author[Jan Hendrik Metzen]{Jan Hendrik Metzen}
\institute[]{University Bremen, Robotics Research Group \\ Robert Bosch GmbH, Corporate Research CR/AEY2}
\date[06/19/16]{ICML, June 19, 2016}

@@

\frame{\titlepage
       \bibliographystyle{abbrv}
       \nobibliography{../literature}}

\begin{frame}{Motivation: Bayesian optimization}

@@

\begin{frame}{Bayesian Optimization}
\textbf{Bayesian optimization}\footcite{shahriari_taking_2016}:
\begin{itemize}
\item black-box optimization problems: $\mathbf{x} = \arg\min_{\mathbf{x} \in \mathcal{X}} f(x)$ of some function $f: \mathcal{X} \to \mathbb{R}$ on some bounded set $\mathcal{X} \subset \mathbb{R}^D$.
\item \emph{probabilistic model} $p(f)$ for $f(\mathbf{x})$, typically a Gaussian process (GP)
\item based on GP posterior of first $n$ observations $\mathcal{D}_n=\{(\mathbf{x}_i, y_i)\}_{i=1}^n$ and an \emph{acquisition function}: decides on a query point $\mathbf{x}_{n+1}$ where $f$ will be evaluated next
\item recommends $\mathbf{\tilde x}_N$ after $N$ queries (optimum of GP or best query point)
\item objective: minimize \emph{simple regret} 
$R_f(\mathbf{\tilde  x}_N) = f(\mathbf{x}^\star) - f(\mathbf{\tilde  x}_N) = \max_{\mathbf{x}} f(\mathbf{x}) - f(\mathbf{\tilde x}_N)$
\end{itemize}
\end{frame}

\begin{frame}{Entropy Search}
\begin{itemize}
\item Let
   \begin{itemize}
   \item $p^\star(x \vert \mathcal{D}_n)$ denote the posterior distribution (after observing $\mathcal{D}_n$) of the unknown optimizer $\mathbf{x}^\star = \arg\max_{\mathbf{x} \in \mathcal{X}} f(\mathbf{x})$.
   \item $H(\mathbf{x}^\star \vert \mathcal{D}_n)$ denote the differential entropy of $p^\star(x \vert \mathcal{D}_n)$
   \end{itemize}
   \item Entropy Search (ES)\footcite{hennig_entropy_2012} is an information theoretic acquisition fct.:
$$a_{ES}(\mathbf{x}, \mathcal{D}_n) = H(\mathbf{x}^\star \vert \mathcal{D}_n) - \mathbb{E}_{y \vert \mathbf{x}, \mathcal{D}_n}   [H(\mathbf{x}^\star \vert \mathcal{D}_n \cup \{(\mathbf{x}, y))]$$
   \end{itemize}
   \pause

   \vspace*{-.25cm}
   \begin{figure}
   \centering
   \includegraphics[width=0.48\columnwidth]{../pics/regret_illustration_2}
   \visible<3->{
     \includegraphics[width=0.48\columnwidth]{../pics/regret_scatter}
   }
   \vspace*{-.25cm}
   \caption{(Left) GP posterior, probability of maximum $p^\star$, and expected regret ER. (Right) Scatter plot of  position of optimum and regret of $\mathbf{\tilde x} = 1.5$ versus optimum in sample functions.}
   \label{fig:MRS_illustration}
   \end{figure}
\end{frame}

\begin{frame}{Minimum Regret Search}

MRS: acquistion function based directly on external objective of minimizing simple regret
\begin{itemize}
 \item Expected simple regret: $\text{ER}(p)(\mathbf{x}) = \mathbb{E}_{p(f)}[R_f(\mathbf{x})]
= \mathbb{E}_{p(f)}[\max_{\mathbf{x}} f(\mathbf{x}) - f(\mathbf{
x})]$
 \item For fixed GP $p(f)$, $\arg\min_\mathbf{x} \text{ER}(p)(\mathbf{x})$ corresponds to the maximizer of the GP mean
 \item MRS aims at selecting query points s.t. ER is minimized also with respect to resulting $p(f)$
 \item Myopic choice: select next query point s.t. minimum ER is reduced the most (in expectation)
  $$a_{\text{MRS}^{\text{point}}}(\mathbf{x}^q)
  = \min_{\mathbf{\tilde x}}\text{ER}(p_n)(\mathbf{\tilde x})
   - \mathbb{E}_{y \vert p_n(f), \mathbf{x}^q}[\min_{\mathbf{\tilde x}}  \text{ER}(p^{[\mathbf{x}^q, y]}_n)(\mathbf{\tilde x})]$$
 \item MRS additionally also accounts for uncertainty in minimizer $\mathbf{\tilde x}$ of $\text{ER}$ 
  $$a_{\text{MRS}}(\mathbf{x}^q)
    = \mathbb{E}_{\mathbf{\tilde x} \sim p^\star_n}[\text{ER}(p_n)(\mathbf{\tilde x})] 
     - \mathbb{E}_{y \vert p_n(f), \mathbf{x}^q}[
      \mathbb{E}_{\mathbf{\tilde x} \sim p^\star_{\mathcal{D}_n \cup \{(\mathbf{x}^q, y)\}}}[
       \text{ER}(p^{[\mathbf{x}^q, y]}_n)(\mathbf{\tilde x})]]$$
\end{itemize}

\begin{frame}{Minimum Regret Search Illustration}
\begin{figure}
\centering
\includegraphics[width=0.48\columnwidth]{../pics/acq_comparison}
\end{figure}

\begin{itemize}
\item MRS$^\text{point}$ tends to query in areas which are risky for current recommendation (large simple regret possible)
\item MRS is more smooth than MRS$^\text{point}$ since it accounts for uncertainty in recommendation
\end{itemize}

\begin{frame}{Experimental Setup}
Synthetic Single-Task Benchmark:
\begin{itemize}
 \item Target functions sampled from a generative model on $\mathcal{X} = [0, 1]^2$ 
 \item In practice: 
   \begin{itemize} 
   \item sample $250$ pairs $(\mathbf{x}, f(\mathbf{x}))$ from function $f \sim GP(f)$
   \item fit another GP to these pairs
   \item use resulting posterior mean as target function
   \end{itemize}
 \item Gaussian noise with standard deviation $\sigma =10^{-3}$ is added to each observation
 \item Probabilistic surrogate model used in BO:\\ GP with isotropic RBF kernel of length scale $l = 0.1$ and unit signal variance
 \pause
 \item Generative model
    \begin{itemize} 
    \item without model mismatch: GP with isotropic RBF kernel ($l = 0.1$ and unit signal variance)
    \item with model mismatch:  GP with isotropic rational quadratic kernel ($l = 0.1$, $\alpha=1.0$ and unit signal variance)
    \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}{Results: No model-mismatch}
\begin{figure}
\centering
\includegraphics[width=.7\columnwidth]{../pics/empirical_comparison} \\
\includegraphics[width=.7\columnwidth]{../pics/hist}
\end{figure}
\end{frame}

\begin{frame}{Results: }
\begin{figure}
\centering
\includegraphics[width=.8\textwidth]{../pics/es_analysis}
\label{fig:es_analysis}
\end{figure}

Acquisition functions on a target function at $N=100$ and 25 representer points; darker areas correspond to larger values. ES focuses on sampling in areas with high density of $p^\star$ (many representer points), while MRS focuses on unexplored areas that are populated by representer points (non-zero $p^\star$).

\begin{frame}{Results: Model-mismatch}
\begin{figure}
\centering
\includegraphics[width=.7\textwidth]{../pics/empirical_comparison_mm} \\
\includegraphics[width=.7\textwidth]{../pics/hist_mm}
\end{figure}
\end{frame}

\begin{frame}{Outlook}

  \begin{itemize}
    \item Additional content of the paper:
    \begin{itemize}
      \item Multi-task Minimum regret search
      \item results on simulated robotic task
    \end{itemize}
    \pause
    \item Future work:
    \begin{itemize}
      \item treatment of GP hyperparameters
      \item more efficient approximation techniques for MRS
    \end{itemize}
    \pause
    \item Source code available at: \url{https://github.com/jmetzen/bayesian_optimization}
  \end{itemize}

  \vspace*{1.5cm}
  \pause
  \begin{center}
   Thank you for your attention! \\
   Do you have questions, comments, or ideas?
   \end{center}
